---
title: "tidyなデータ"
subtitle: "Kanazawa.R #1"
date: 2024-06-29
author: "伊東宏樹"
lang: ja
format:
  revealjs:
    theme: [dark, custom.scss]
    code-copy: true
    template-partials:
      - title-slide.html
    fig-width: 5
    fig-height: 3
execute:
  echo: true
slide-number: true
editor: visual
embed-resources: true
---

## 内容

-   Tidy data（整然データ）

-   tidyverse

    -   tidyr

    -   dplyr

## Tidy data

When you click the **Render** button a document will be generated that includes:

-   Content authored with markdown
-   Output from executable code

## tidyverse

```{r}
#| label: tidyverse

library(tidyverse)
```

## tidyr

-   データの変形

    -   -\> long

    -   -\> wide

## データ

世界銀行による各国の人口データ

```{r}
#| label: data

data(world_bank_pop)
head(world_bank_pop)
```

## こうしたい

::: {style="font-size: 66%;"}
| country | year | [SP.URB.TOTL]{style="font-size: 80%;"} | [SP.URB.GROW]{style="font-size: 80%;"} | [SP.POP.TOTL]{style="font-size: 80%;"} | [SP.POP.GROW]{style="font-size: 80%;"} |
|------------|-----------:|-----------:|-----------:|-----------:|-----------:|
| ABW     | 2000 |                                  41625 |                                   1.66 |                                  89101 |                                   2.54 |
| ABW     | 2001 |                                  42025 |                                  0.956 |                                  90691 |                                   1.77 |
|   :     |  :   |                                    :   |                                    :   |                                    :   |                                    :   |
| ABW     | 2017 |                                  45648 |                                  0.772 |                                 105439 |                                  0.537 |
| AFE     | 2000 |                              115551653 |                                   3.60 |                              401600588 |                                   2.58 |
|   :     |  :   |                                    :   |                                    :   |                                    :   |                                    :   |
| ZWE     | 2017 |                                4755312 |                                   1.86 |                               14751101 |                                   2.04 |
:::

## pivot_longer

```{r}
#| label: pivot_longer

pop_long <- world_bank_pop |>
  pivot_longer(cols = `2000`:`2017`, names_to = "year")
pop_long
```

## pivot_wider

```{r}
#| label: pivot_wider

pop_tidy <- pop_long |>
  pivot_wider(names_from = indicator)
pop_tidy
```

## dplyr

## mutate

```{r}
#| label: mutate

pop_tidy |>
  mutate(pop_non_urb = SP.POP.TOTL - SP.URB.TOTL)
```

## transmute

```{r}
#| label: transmute

pop_tidy |>
  transmute(country, year, pop_non_urb = SP.POP.TOTL - SP.URB.TOTL)
```

## filter

```{r}
#| label: filter

pop_tidy |>
  filter(year == 2017, SP.POP.TOTL >= 1e+8)
```

## select

```{r}
#| label: select

pop_tidy |>
  select(country, year, SP.POP.TOTL)
```

## arrange

```{r}
#| label: arrange

pop_tidy |>
  arrange(desc(SP.POP.TOTL))
```

## summarise

```{r}
#| label: summarise

pop_tidy |>
  group_by(country) |>
  summarise(mean_pop = mean(SP.POP.TOTL))
```

## \*\_join

## 参考情報
